Capstone Movielens Report

### 2025-05-28 reports/capstone-movielens.site/02.7.UMGYE-model.Rmd
The model mathematical description completed.
fix for UMGYE Model; dev: function described: calc_day_general_effect, calc_day_general_effect described.

This is a helper function that calculates the _date general effect_ of the user ratings and is intended to be used in other functions and scripts (for instance in [train_UMGY_effect](#func.train_UMGY_effect)) described below.

@@

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L2) function is available in the [UMGYE Model Support Functions](#L1) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L2) script on _GitHub_.
:::

@@

A data structure containg information about _Date General Effect_.

@@

This is a helper function that calculates the _date general effect_ of the user ratings using the _5-Fold Cross Validation_ method and is intended to be used in other functions and scripts (for instance in [train_UMGY_effect.cv](#func.train_UMGY_effect.cv)) described below.

@@

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L30) function is available in the [UMGYE Model Support Functions](#L1) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L30) script on _GitHub_.
:::

@@

::: {.noteblock data-latex=""}
Here we use the function call [union_cv_results](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L432) to aggregate the _5-Fold Cross Validation_ method results (for details, see the [union_cv_results](#func.union_cv_results) function description in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report).
:::

@@

A data structure containg information about _Date General Effect_ as result of _5-Fold Cross Validation_.

@@

This is a helper function that calculates the _UMGY effect_ of the user ratings and is intended to be used in other functions and scripts (for instance in [train_UMGY_effect](#func.train_UMGY_effect)) described below.

@@

A data structure containg information about _User+Movie+Genre+Year (UMGY) Effect_.

@@

where $n_r$ is the number of ratings made in year $r$.

>>>>

This is a helper function that calculates the _date general effect_ of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGY_effect](#func.train_UMGY_effect)) described below.

@@

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L2) function is available in the [UMGYE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L1) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L2) script on _GitHub_.
:::

@@

A data structure containing information about _Date General Effect_.

@@

This is a helper function that calculates the _date general effect_ of the user ratings using the _5-Fold Cross Validation_ method and is intended to be used in other functions and scripts (such as [train_UMGY_effect.cv](#func.train_UMGY_effect.cv)) described below.

@@

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L30) function is available in the [UMGYE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L1) section of the [UMGY-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGY-effect.functions.R#L30) script on _GitHub_.
:::

@@

::: {.noteblock data-latex=""}
Here we use the function call [union_cv_results](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L432) to aggregate the _5-Fold Cross Validation_ results (for details, see the [union_cv_results](#func.union_cv_results) function description in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report).
:::

@@

A data structure containing information about _Date General Effect_ as result of _5-Fold Cross Validation_.

@@

This is a helper function that calculates the _UMGY effect_ of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGY_effect](#func.train_UMGY_effect)) described below.

@@

A data structure containing information about _User+Movie+Genre+Year (UMGY) Effect_.

@@

where $n_v$ is the number of ratings made in year $v$.

### 2025-05-28 reports/capstone-movielens.site/02.8-0.UMGYDE-model.Rmd
fix for UMGYE Model; dev: function described: calc_day_general_effect, calc_day_general_effect described.

### Day Effect Analysis

::: {.noteblock data-latex=""}
The _Year Effect_ visualization and analysis code in this section are cited from the article [Movie Recommendation System using R - BEST](https://www.kaggle.com/code/amirmotefaker/movie-recommendation-system-using-r-best/notebook) mentioned earlier in this report[@MRS-R-BEST].
:::

@@

If we define $d_{i,j}$ as the _number of days since the earliest record_ in the `edx` dataset for movie $j$ rated by user $i$, then the formula \@ref(eq:UMGYE-model) describing the _UMGYE Model_, for the current model, takes the form:

@@

@@

This is a helper function that calculates the _UMGYD effect_ of the user ratings and is intended to be used in other functions and scripts (for instance in [train_UMGYD_effect](#func.train_UMGYD_effect)) described below.

@@

  - [calc_date_general_effect](#func.calc_date_general_effect);
  - [calc_date_general_effect.cv](#func.calc_date_general_effect.cv).

@@

A data structure containg information about _User+Movie+Genre+Year (UMGYD) Effect_.

@@

  train_set |>
    calc_date_general_effect(lambda) |>
    calc_UMGYD_effect()
}

@@

  calc_date_general_effect.cv(lambda) |> calc_UMGYD_effect()  
}
```

@@

We have already explained the idea of _Linear Model Regularization_ in the [UME Model Regularization] section above. We have also seen how the formula \@ref(eq:ME-penalty) for adding a penalty to the _UME Model_ is transformed into the formula \@ref(eq:GE-penalty) for the _UMGE Model_ and then into the formula \@ref(eq:GYE-penalty) for the _UMGYE Model_. For the current model, this formula takes the form:

>>>>

@@

If we define $d_{i,j}$ as the _number of days since the earliest record_ in the `edx` dataset for movie $j$ rated by user $i$, then the formula \@ref(eq:UMGYDE-model) describing the _UMGYDE Model_, for the current model, takes the form:

@@

#### [calc_day_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) Function {#func.calc_day_general_effect}
\

This is a helper function that calculates the _day general effect_ of user ratings on a _given day since the earliest record_ in the `edx` dataset and is intended to be used in other functions and scripts (such as [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect)) described below.

The following is a simplified version of the function's source code:
```{r eval=FALSE}
calc_day_general_effect <- function(train_set, lambda = 0){
  train_set |> 
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    left_join(rglr.UMGY_effect, by='year') |>
    mutate(resid = rating - (mu + a + b + g + ye)) |>
    group_by(days) |>
    summarise(de = mean_reg(resid, lambda), 
              year = mean(year))
}
```

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_day_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) function is available in the [UMGYDE Model Support Functions]((https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) script on _GitHub_.
:::

A data structure containing information about _Day General Effect_.

#### [calc_day_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L31) Function {#func.calc_day_general_effect.cv}
\

This is a helper function that calculates the _day general effect_ of user ratings on a _given day since the earliest record_ in the `edx` dataset using the _5-Fold Cross Validation_ and is intended to be used in other functions and scripts (such as [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv)) described below.

##### Souurce Code
\

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_day_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L31) function is available in the [UMGYDE Model Support Functions](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L30) script on _GitHub_.
:::

The following is a simplified version of the function's source code:
```{r eval=FALSE}
calc_day_general_effect.cv <- function(lambda = 0){
  gday_effect_ls <- lapply(edx_CV,  function(cv_fold_dat){
    cv_fold_dat$train_set |> calc_day_general_effect(lambda)
  })

  gday_effect_united <- union_cv_results(gday_effect_ls)

  gday_effect_united |>
    group_by(days) |>
    summarise(de = mean(de), year = mean(year))
}
```

::: {.noteblock data-latex=""}
Here we use the function call [union_cv_results](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L432) to aggregate the _5-Fold Cross Validation_ results (for details, see the [union_cv_results](#func.union_cv_results) function description in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report).
:::

A data structure containing information about _Date General Effect_ as result of _5-Fold Cross Validation_.

@@

This is a helper function that calculates the _UMGYD effect_ of the user ratings and is intended to be used in other functions and scripts (such as [train_UMGYD_effect](#func.train_UMGYD_effect)) described below.

@@

  - [calc_day_general_effect](#func.calc_day_general_effect);
  - [calc_day_general_effect.cv](#func.calc_day_general_effect.cv).

@@

A data structure containing information about _User+Movie+Genre+Year (UMGYD) Effect_.

@@

  train_set |>
    calc_day_general_effect(lambda) |>
    calc_UMGYD_effect()
}

@@

  calc_day_general_effect.cv(lambda) |> calc_UMGYD_effect()  
}
```

@@

We have already explained the idea of _Linear Model Regularization_ in the [UME Model Regularization] section above. We have also seen how the formula \@ref(eq:ME-penalty) for adding a penalty to the _UME Model_ is transformed into the formula \@ref(eq:GE-penalty) for the _UMGE Model_ and then into the formula \@ref(eq:GYE-penalty) for the _UMGYDE Model_. For the current model, this formula takes the form:
